Computing in the Field of Rationals

### 2012-06-25 rational_ext.tex
_

Andrzej Filinski's Masters thesis
\cite{Filinski:1989:DCI:648332.755574} suggested a remarkable symmetry
underlying computation in the \lcal. Filinski showed that values are
dual to continuations, functions are dual to delimited
continuations. Filinski also shows a De Morgan style duality by
treating pairs of values as dual to sums (choice) of continuation and
sums of values as dual to pairs of continuations. He, furthermore,
introduces a symmetric extension of the $\lambda$-calculus in which
call-by-value is dual to call-by-name.

Symmetry between the value fragment of the language and the
continuation fragment was refined by Curien and Herbelin
\cite{Curien:2000} who introduced the sequent-style
$\mu\tilde{\mu}$-calculus, that exhibits symmetries between values and
continuations and between call-by-value and call-by-name. The calculus
includes the type $A-B$ which is the dual of implication, i.e., a
value of type $A-B$ is a context expecting a function of type $A
\rightarrow B$. Alternatively a value of type $A-B$ is also explained
as a \emph{pair} consisting of a value of type $A$ and a continuation
of type $B$. This is to be contrasted with our interpretation of a
value of that type as \emph{either} a value of $A$ or a demand for a
value of type $B$. The duality between call-by-name and call-by-value
was further investigated by Selinger using control
categories~\cite{Selinger:2001:CCD:966910.966911}. The
$\mu\tilde{\mu}$-calculus was further analyzed and extended by
Wadler~\cite{Wadler:2003,DBLP:conf/rta/Wadler05}. The extension gives
no interpretation to the subtraction connective and like the original
symmetric calculus of Filinski, introduces a duality that relates sums
to products and vice-versa.

@@

The language {{Pi}} however lacks a natural representation for
first-class functions due to first-order nature of all of its
values. In the light of the work of duality of computation it was felt
that ``right model'' for expressing functions in {{Pi}} would also
express symmetries similar to those in the calculi of Filinski, Curien
and Herbelin and Wadler. The technical goal of this paper is to
present such a symmetric dual extension in the setting of {{Pi}}.

>>>>

Andrzej Filinski's Masters thesis
\cite{Filinski:1989:DCI:648332.755574} suggested a remarkable symmetry
underlying computation in the \lcal. Filinski showed that values are
dual to continuations, functions are dual to delimited continuations
and a De Morgan style duality by treating pairs of values as dual to
sums (choice) of continuation and sums of values as dual to pairs of
continuations. He, furthermore, introduced a symmetric extension of
the $\lambda$-calculus in which call-by-value is dual to call-by-name.

Symmetry between the value fragment and the continuation fragment was
refined by Curien and Herbelin \cite{Curien:2000} who introduced the
sequent-style $\mu\tilde{\mu}$-calculus, that exhibits symmetries
between values and continuations and between call-by-value and
call-by-name. The calculus includes the type $A-B$ which is the dual
of implication, i.e., a value of type $A-B$ is a context expecting a
function of type $A \rightarrow B$. Alternatively a value of type
$A-B$ is also explained as a \emph{pair} consisting of a value of type
$A$ and a continuation of type $B$. This is to be contrasted with our
interpretation of a value of that type as \emph{either} a value of $A$
or a demand for a value of type $B$. The duality between call-by-name
and call-by-value was further investigated by Selinger using control
categories~\cite{Selinger:2001:CCD:966910.966911}. The
$\mu\tilde{\mu}$-calculus was further analyzed and extended by
Wadler~\cite{Wadler:2003,DBLP:conf/rta/Wadler05}. The extension gives
no interpretation to the subtraction connective and like the original
symmetric calculus of Filinski, introduces a duality that relates sums
to products and vice-versa.

@@

The language {{Pi}} however lacks a natural representation for
first-class functions due to the first-order nature of its values. In
the light of the work of duality of computation it was felt that
``right model'' for expressing functions in {{Pi}} would also express
symmetries similar to those in the calculi of Filinski, Curien and
Herbelin and Wadler. The technical goal of this paper is to present
such a symmetric dual extension of {{Pi}}.
